---
title: "Figure5+6-pHManualvsSensor"
---
## Loading Packages
- Package files and their corresponding manuals can be found in the '02_Packages' 

```{r include=TRUE, message=FALSE, echo=TRUE}
library(readxl) #needed for loading xlsx file
library(tidyverse) #contains ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats packages 
library(rstatix) #needed for tukey_hsd function
library(ggpubr) #requires tidyverse to be loaded first, needed for graphs (e.g. ggboxplot function, and ggpar
```

## Loading Sensor Data1 - Manually extracted pH - For Fig 5a and Fig6a
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
ManualData<-read_excel("TableS1-pH+mV-Manual+Sensor.xlsx",sheet = "Manual-pH", col_types = c("numeric","numeric","text","numeric","text", "text")) # loading excel, definning which sheet is loaded and define the type of column each is (have to set manually) - important later for stats
ManualData$Cultivar <- factor(ManualData$Cultivar, levels = c("WM-12", "WM-1","G122","Eldorado","ICA Bunsi","Beryl", "Othello","Montrose")) #order the Cultivars (for graph later on)
ManualData$TimePoint <- factor(ManualData$TimePoint, levels = c("Pre treatment", "Post treatment")) #order the treatments pre and post treatment (for graph later on)
```

# Fig5a - Manual Extracted pH - Wounded (Block Number = 4, Note = n=1 vs n=3 in Blocks 1-3, which are inoculated)
## Splitting Data
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
ManualDataWound <- ManualData%>% filter(Block == 4) #Create a new table from ManualData that just has Block 4 (Wounded plants)
ManualData<- ManualData %>% filter(Block < 4) #removes Block 4 from ManualData - will be used for inoculated plots in Fig6a
```

## Stats- Two-way ANOVA - Ss response and TimePoint (i.e. Pre-and post treatment). Data showed in Table S2.
```{r echo=FALSE}
summary(aov(pH ~ Ss_Response + TimePoint + Ss_Response*TimePoint, data = ManualDataWound))
```
## Post-hoc Tests for calculating p values when comparing pre and post treatment within each SS response group (resistant, intermediate tolerant, susceptible) - then manually add p values to Fig5a in ppt.
```{r echo=FALSE}
Fig5aposthoc<-ManualDataWound %>% group_by(Ss_Response) %>% tukey_hsd(pH ~ TimePoint)
Fig5aposthoc
```

## Generating Figure5a
```{r echo=FALSE}
Fig5a<-ggplot(ManualDataWound,aes(x=Cultivar,y=pH, fill=TimePoint)) + # Fill=TimePoint means the colour is based on the TimePoint column i.e. post or pre treatment 
geom_point(aes(color = TimePoint), size = 3, position = position_dodge(width=0.3)) + #adds data points to plot
xlab("Cultivar") + #Set x axis label
theme_bw() + #Set theme of graph
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ #remove previous grid lines that are added with theme_bw 
theme(panel.border = element_blank(), panel.grid.major = element_blank(), axis.line = element_line(colour = "black")) + #remove previous grid lines that are added with theme_bw
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1, color="black")) + # x axis font size
theme(axis.text.y = element_text(size=12, color="black")) + # y axis font size
theme(plot.title = element_text(color="black", size=12)) + # change title font size
xlab("") + #remove the x axis label (Cultivar)
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5) #adding in a dashed line
Fig5a <- Fig5a + theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="white", linetype="blank")) #adding grid lines to graph, minor grids (0.5) are blank to remove them
Fig5a <- Fig5a + ggtitle("Manual - Leaf Apoplastic Fluid") + #Defining title
theme(plot.title = element_text(hjust = 0.5, vjust=2)) #Centering title
Fig5a <- Fig5a + theme(axis.title.x = element_text(size=12)) #changing x axis label font size
Fig5a <- Fig5a + theme(axis.title.y = element_text(size=12)) #changing y axis label size
Fig5a<-ggpar(Fig5a, legend.title = "") # Remove legend title which was 'TimePoint'
Fig5a<-ggpar(Fig5a, legend = "bottom") # Legend at bottom
Fig5a
```


# Figure5b - Sensor pH converted - Wounded (Block Number = 4, Note = n=1 vs n=3 in Blocks 1-3, which are inoculated)
## Loading SensorData2 - Sensor coverted pH - For Figure5b and Figure6b
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
SensorData<-read_excel("TableS1-pH+mV-Manual+Sensor.xlsx",sheet = "Sensor-pH", col_types = c("numeric","numeric","text","numeric","text", "text")) # loading excel, definning which sheet is loaded and define the type of column each is (have to set manually) - important later for stats
SensorData$Cultivar <- factor(SensorData$Cultivar, levels = c("WM-12", "WM-1","G122","Eldorado","ICA Bunsi","Beryl", "Othello","Montrose")) #order the Cultivars (for graph later on)
```

## Splitting Data
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
SensorDataWound <- SensorData %>% filter(Block == 4) #Create a new table from SensorData that just has Block 4 (Wound treated)
SensorData <- SensorData %>% filter(Block < 4) #removes Block 4 from SensorData  - will be used for inoculated plots in Fig6b
```

## Stats- Two-way ANOVA - Cultivar and TimePoint (i.e. Pre-and post treatment) for Figure5b. Data showed in Table S3.
```{r echo=FALSE}
summary(aov(pH ~ Cultivar + TimePoint + Cultivar*TimePoint, data = SensorDataWound))
```

## Post-hoc Tests for calculating individual p values when comparing pre and post treatment within each Cultivar - then manually add p values to Figure5b in ppt.
```{r echo=FALSE}
Fig5bposthoc<-SensorDataWound %>% group_by(Cultivar) %>% tukey_hsd(pH ~ TimePoint)
Fig5bposthoc
```

#Figure5b - Sensor pH Wounded
```{r message=FALSE, warning=FALSE}
SensorDataWound$TimePoint <- factor(SensorDataWound$TimePoint, levels = c("Pre treatment", "Post treatment")) #ensure that pre and then post treatment appears in that order on plot 2
Fig5b<-ggboxplot(SensorDataWound,"Cultivar", "pH", add="jitter",color="TimePoint", title="Sensor - Leaf Surface mV Converted to pH")+
theme(plot.title = element_text(hjust = 0.5, vjust = 2)) + #Move title to middle, move up a bit
theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="white", linetype="blank")) + #adding grid lines to graph, minor grids (0.5) are blank to remove them
theme(plot.title = element_text(color="black", size=12)) + # change title font size
xlab("") + #remove the x axis label (Cultivar)
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1,color="black")) + # x axis font size
theme(axis.text.y = element_text(size=12)) + # y axis font size
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5,color="black") + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5,color="black") #adding in a dashed line
Fig5b <- Fig5b + theme(axis.title.x = element_text(size=12)) #changing x axis label font size
Fig5b <- Fig5b + theme(axis.title.y = element_text(size=12)) #changing y axis label size
Fig5b<-ggpar(Fig5b, legend = "bottom") # Legend at bottom
Fig5b<-ggpar(Fig5b, legend.title = "") # Remove legend title which was 'TimePoint'
Fig5b
```

#Combining Fig5a and Fig5b to make Fig5
```{r}
Fig5 <- ggarrange(Fig5a, Fig5b, ncol = 2, nrow = 1, labels = "auto", legend = "bottom", widths = c(1.4,2)) #width squashes Panel A graph
Fig5 <- annotate_figure(Fig5, top = text_grob("Wounding Treatment", color = "black", face = "bold", size = 12)) #Adding a main title above the panels
Fig5
```

#Saving Figure5
```{r}
#Saving the plot as a tiff and pdf
ggsave(plot=Fig5, file="Figure5-pH-Wounded.tiff", units="cm", width=18, height=12, dpi=600, compression='lzw')
ggsave(plot=Fig5, file="Fig5-pH-Wounded.pdf", units="cm", width=18, height=12, dpi=600)
```




# Figure6a - Manual Extracted pH - Inoculated (Blocks= 1-3, Note = n=3)
## Stats- Two-way ANOVA - Cultivar and TimePoint (i.e. Pre-and post treatment) for Fig6a. Data showed in Sup. TableS5
```{r echo=FALSE}
summary(aov(pH ~ Cultivar + Block + TimePoint + Cultivar*TimePoint, data = ManualData))
```
## Post-hoc Tests:
## Post-hoc Tests for calculating individual p values when comparing pre and post treatment within each Cultivar - then manually add p values to Fig6a in ppt
```{r echo=FALSE}
Fig6aposthoc<-ManualData %>% group_by(Cultivar) %>% tukey_hsd(pH ~ TimePoint)
Fig6aposthoc
```

#Fig.6a inoculated
```{r message=FALSE, warning=FALSE}
Fig6a<-ggboxplot(ManualData,"Cultivar", "pH", add="jitter",color="TimePoint", title="Manual - Leaf Apoplastic Fluid", ylim=c(6.2,7.3)) +
theme(plot.title = element_text(hjust = 0.5, vjust=2)) + #Move title to middle 
theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="white", linetype="blank")) + #adding grid lines to graph, minor grids (0.5) are blank to remove them
theme(plot.title = element_text(color="black", size=12)) + # change title font size
xlab("") + #remove the x axis label (Cultivar)
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1,color="black")) + # x axis font size
theme(axis.text.y = element_text(size=12)) + # y axis font size
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5,color="black") + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5,color="black") #adding in a dashed line
Fig6a<-ggpar(Fig6a, legend = "bottom") # Legend at bottom
Fig6a<-ggpar(Fig6a, legend.title = "") # Remove legend title which was 'TimePoint'
Fig6a
```


# Fig6b - Sensor Converted pH - Inoculated (Blocks= 1-3, Note = n=3)
## Stats - Two-way ANOVA - Cultivar and TimePoint (i.e. Pre-and post treatment) for Plot 3 - presented in Sup. Table S6
```{r echo=FALSE}
summary(aov(pH ~ Cultivar + Block + TimePoint + Cultivar*TimePoint, data = SensorData))
#Cultivar*TimePoint is asking if there is significant impacting if you look take into account TimePoint (i.e. pre or post) which asking if the Cultivar pH is significantly different in general (not looking at individual Cultivars at this point - not until posthoc test)
```
## Post-hoc Tests:
## Post-hoc Tests for calculating individual p values when comparing pre and post treatment within each Cultivar - then manually add p values to Fig6b in ppt
```{r echo=FALSE}
Fig6bposthoc<-SensorData %>% group_by(Cultivar) %>% tukey_hsd(pH ~ TimePoint)
Fig6bposthoc
```

#Fig6b - Sensor pH Inoculated
```{r message=FALSE, warning=FALSE}
SensorData$TimePoint <- factor(SensorData$TimePoint, levels = c("Pre treatment", "Post treatment")) #ensure that pre and then post treatment appears in that order on plot 2
Fig6b<-ggboxplot(SensorData,"Cultivar", "pH", add="jitter",color="TimePoint", title="Sensor - Leaf Surface mV Converted to pH")+
theme(plot.title = element_text(hjust = 0.5, vjust=2)) + #Move title to middle 
theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="white", linetype="blank"))+ #adding grid lines to graph, minor grids (0.5) are blank to remove them
theme(plot.title = element_text(color="black", size=12)) + # change title font size
xlab("") + #remove the x axis label (Cultivar)
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1,color="black")) + # x axis font size
theme(axis.text.y = element_text(size=12)) + # y axis font size
theme(axis.title.x = element_text(size=11)) + #changing x axis label font size
theme(axis.title.y = element_text(size=11))  +#changing y axis label font size
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5,color="black") + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5,color="black") #adding in a dashed line
Fig6b<-ggpar(Fig6b, legend = "bottom") # Legend at bottom
Fig6b<-ggpar(Fig6b, legend.title = "") # Remove legend title which was 'TimePoint'
Fig6b
```

#Combing Fig6a and Fig6b to make Figure6
```{r message=FALSE, warning=FALSE}
Fig6 <- ggarrange(Fig6a, Fig6b, ncol = 2, nrow = 1, labels = "auto", legend = "bottom", widths = c(1.4,2)) #width squashes Panel A graph
Fig6 <- annotate_figure(Fig6, top = text_grob("Inoculation Treatment (Wounding + Ss)", color = "black", face = "bold", size = 12)) #Adding a main title above the panels
Fig6
```

#Saving Figure6
```{r eval=FALSE, include=FALSE}
#Saving the plot as a tiff and pdf
ggsave(plot=Fig6, file="Figure6-pH-Inoculated.tiff", units="cm", width=18, height=12, dpi=600, compression='lzw')
ggsave(plot=Fig6, file="Figure6-pH-Inoculated.pdf", units="cm", width=18, height=12, dpi=600)
```





#Analysis of difference in change in pH detected using the Manual pH or Sensor pH Approach
## Loading pH Change data
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
CombinedData<-read_excel("TableS1-pH+mV-Manual+Sensor.xlsx",sheet = "Combined-pH-change", col_types = c("numeric","numeric","text","numeric","text", "text")) # loading excel, definning which sheet is 
CombinedDataWound<- CombinedData%>% filter(Block == 4) #Create a new table from CombinedData that just has Block 4 (wounding)
CombinedData<- CombinedData %>% filter(Block < 4) #removes Block 4 from CombinedData
```

#Wound Treated
## Stats- Two-way ANOVA of the Change in pH (Post-Pre) 'pH_change' vs 'Cultivar' vs 'TimePoint'(Pre or Post) vs 'Measurement_Type' i.e. either 'Manual' Leaf apoplastic fluid or 'Sensor' leaf surface mV converted to pH 'Measurement_Type'. Presented in Table S3.
#Wound Treated plants
```{r echo=FALSE}
summary(aov(pH_change ~ Cultivar + Measurement_Type +Cultivar/Measurement_Type, data = CombinedDataWound)) 
```

## Post-hoc Tests for calculating individual p values when comparing pre and post treatment within each Cultivar. Presented in Table S4.
```{r echo=FALSE}
WoundpHchangeposthoc<-CombinedDataWound %>% group_by(Cultivar) %>% tukey_hsd(pH_change ~ Measurement_Type) 
WoundpHchangeposthoc
```

#Inoculation Treated
## Stats- Two-way ANOVA of the Change in pH (Post-Pre) 'pH_change' vs 'Cultivar' vs 'TimePoint'(Pre or Post) vs 'Measurement_Type' i.e. either 'Manual' Leaf apoplastic fluid or 'Sensor' leaf surface mV converted to pH 'Measurement_Type'.Presented in Table S7.
#Inoculated plants
```{r echo=FALSE}
summary(aov(pH_change ~ Cultivar + Block + Measurement_Type +Cultivar/Measurement_Type, data = CombinedData)) 
```

##  Post-hoc Tests for calculating individual p values when comparing pre and post treatment within each Cultivar. Presented in Table S8.
```{r echo=FALSE}
InoculatedpHchangeposthoc<-CombinedData %>% group_by(Cultivar) %>% tukey_hsd(pH_change ~ Measurement_Type)
InoculatedpHchangeposthoc
```
