---
title: "Fig.4-Hysteresis of carbon sensors-RM"
---

## Loading Packages
- Package files and their corresponding manuals can be found in the '02_Packages' folder in the current working directory.
```{r include=TRUE, error=FALSE,warning=FALSE,echo=FALSE}
#Note, message=FALSE since a number of these packages have the same named functions.
library(readxl) #needed for loading xlsx file
library(dplyr) #Has select function
library(ggplot2) #Required for graphing
library(ggpubr) #Has gg arrange function
```

## Loading Data
- Excel file with raw data is in the 'Analysis' folder in the current working directory, which can be double checked by typing in 'getwd()' in the console.
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
ph <- read_excel("Fig4-Hysteresis.xlsx",sheet = 2, col_types = c("numeric", "numeric","numeric","numeric","numeric")) #load data and define column type, all five columns are numeric
```

## Filtering 'ph' by sensor type and inverse transformation of Carbon Sensor 2
```{r}
sensor1 <- ph %>% select(1,2,3) #select columns 1, 2, and 3 (where sensor 1 mV values are)
sensor2 <- ph %>% select(1,2,4) #select columns 1, 2, and 4 (where sensor 2 mV values are)
sensor3 <- ph %>% select(1,2,5) #select columns 1, 2, and 5 (where sensor 3 mV values are

sensor2[4] <- (-sensor2[3]+max(sensor2[3]))/2  #Carbon sensor 2 had reverse wiring so - inverse transformed mV to create a new column sensor2[4]. Formaula:(- negative mV + max mV) / 2

colnames(sensor1) <- c("Time","pH","mV") #set column names
colnames(sensor2) <- c("Time","pH","mV", "mV.transform") #set column names
colnames(sensor3) <- c("Time","pH","mV") #set column names
```


## Calculating Hysteresis Values for each sensor type
```{r}
Sensor1_First<-filter(sensor1, pH == 3.52) # Filter ph table with rows with pH:3.52 (171-236s) 
Sensor1_Second<-filter(sensor1, pH == 3.94) # Filter ph table with rows with pH 3.94 (Time 693-746s)

Sensor2_First<-filter(sensor2, pH == 3.52) # Filter sensor 2 table with rows with pH:3.52 (171-236s) 
Sensor2_Second<-filter(sensor2, pH == 3.94) # Filter sensor 2 table with rows with pH 3.94 (693-746s)

Sensor3_First<-filter(sensor3, pH == 3.52) # Filter ph table with rows with pH:3.52 (171-236s) 
Sensor3_Second<-filter(sensor3, pH == 3.94) # Filter ph table with rows with pH 3.94 (Time 693-746s)

Hysteresis_Sensor1<-(abs((mean(Sensor1_First$mV)) - (mean(Sensor1_Second$mV)))) #For Sensor 1 (Sensor.1 column) output the difference in mean mV between pH3.52 and pH3.94 and use the abs function to transform negative to positive value if needed.
Hysteresis_Sensor1<-round(Hysteresis_Sensor1, digits=2) #Round to 2 decimal points

Hysteresis_Sensor2<-(abs((mean(Sensor2_First$mV.transform)) - (mean(Sensor2_Second$mV.transform)))) #For Sensor 2 (mV.transform) output the difference in mean mV between pH3.52 and pH3.74 and use the abs function to transform negative to positive value if needed.
Hysteresis_Sensor2<-round(Hysteresis_Sensor2, digits=2) #Round to 2 decimal points

Hysteresis_Sensor3<-(abs((mean(Sensor3_First$mV)) - (mean(Sensor3_Second$mV)))) #For Sensor 3 (Sensor.3 column) output the difference in mean mV between pH3.52 and pH3.74 and use the abs function to transform negative to positive value if needed.
Hysteresis_Sensor3<-round(Hysteresis_Sensor3, digits=2) #Round to 2 decimal points
```


## Graphs
```{r}
carbonsensor1 <- ggplot(sensor1, aes(x=Time, y=mV, color=pH)) +
  geom_line() +
  labs(title="Carbon Sensor 1", x="Time (s)", y = "Millivolt (mV)") + #Setting title and axis
  theme_classic() + #setting theme
  geom_text(aes(x = 740, y = 63, label=paste ("Hysteresis:pH 3.7:", Hysteresis_Sensor1, "mV")), colour="black", size = 3) + #adding in hysteresis value and text
  scale_color_gradient(low = "yellow", high = "darkgreen") + #colour gradient
  geom_vline(xintercept = 171, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 171s (where pH was increased)
  geom_vline(xintercept = 493, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 493s (where pH was decreased)
  theme(plot.title = element_text(color="black", size=8)) + # change title font size
  theme(axis.text.x = element_text(size=8, color="black"))+ # x axis font size and color
  theme(axis.text.y = element_text(size=8, color="black")) # y axis font size and color
  carbonsensor1<-carbonsensor1+theme(axis.title.x = element_text(size=8, color="black")) #changing x axis label font size and color
  carbonsensor1<-carbonsensor1+theme(axis.title.y = element_text(size=8, color="black")) #changing x axis label font size and color

carbonsensor2 <- ggplot(sensor2, aes(x=Time, y=mV.transform, color=pH)) +
  geom_line() + #setting theme
  labs(title="Carbon Sensor 2", x="Time (s)", y = "Millivolt (mV)") + #Setting title and axis
  theme_classic() +
  geom_text(aes(x = 740, y = 73, label=paste ("Hysteresis:pH 3.7:", Hysteresis_Sensor2, "mV")), colour="black", size = 3) + #adding in hysteresis value and text
  scale_color_gradient(low = "yellow", high = "darkgreen") + #colour gradient
  geom_vline(xintercept = 171, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 171s (where pH was increased)
  geom_vline(xintercept = 493, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 493s (where pH was decreased)
  theme(plot.title = element_text(color="black", size=8)) + # change title font size
  theme(axis.text.x = element_text(size=8, color="black"))+ # x axis font size and color
  theme(axis.text.y = element_text(size=8, color="black"))  # y axis font size and color
  carbonsensor2<-carbonsensor2+theme(axis.title.x = element_text(size=8, color="black")) #changing x axis label font size and color
  carbonsensor2<-carbonsensor2+theme(axis.title.y = element_text(size=8, color="black")) #changing x axis label font size and color

carbonsensor3 <- ggplot(sensor3, aes(x=Time, y=mV, color=pH)) +
  geom_line() +
  labs(title="Carbon Sensor 3", x="Time (s)", y = "Millivolt (mV)") + #Setting title and axis
  theme_classic() + #setting theme
  geom_text(aes(x = 740, y = 60, label=paste ("Hysteresis:pH 3.7:", Hysteresis_Sensor3, "mV")), colour="black", size = 3) + #adding in hysteresis value and text
  scale_color_gradient(low = "yellow", high = "darkgreen") + #colour gradient
  geom_vline(xintercept = 171, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 171s (where pH was increased)
  geom_vline(xintercept = 493, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line at 493s (where pH was decreased)
  theme(plot.title = element_text(color="black", size=8)) + # change title font size
  theme(axis.text.x = element_text(size=8, color="black"))+ # x axis font size and color
  theme(axis.text.y = element_text(size=8, color="black")) # y axis font size and color
  carbonsensor3<-carbonsensor3+theme(axis.title.x = element_text(size=8, color="black")) #changing x axis label font size and color
  carbonsensor3<-carbonsensor3+theme(axis.title.y = element_text(size=8, color="black")) #changing x axis label font size and color
```

## Arranging x3 graphs to 1 figure
```{r}
Fig4 <- ggarrange(carbonsensor1, carbonsensor2, carbonsensor3, labels = c("a", "b", "c"), ncol = 1, nrow = 3)
Fig4 <- annotate_figure(Fig4, top = text_grob("Hysteresis of Carbon Sensors 1, 2, and 3", color = "black", face = "bold", size = 8)) #Adding a main title above the panels
Fig4
```
## Saving Fig4
```{r}        
ggsave(file="Fig4-Hysteresis.tiff", plot=Fig4, units="cm", width=18, height=12, compression='lzw')
ggsave(file="Fig4-Hysteresis.pdf", plot=Fig4, units="cm", width=18, height=12)
```

