---
title: "FigureS1 + FigureS2 - Leaf mV"
---
## Loading Packages
- Package files and their corresponding manuals can be found in the '02_Packages' folder

```{r include=TRUE, message=FALSE, echo=TRUE}
library(readxl) #needed for loading xlsx file
library(tidyverse) #contains ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats packages 
library(ggpubr) #requires tidyverse to be loaded first, needed for graphs (e.g. ggboxplot function)
```

## Loading Sensor Data1 - Manually extracted pH
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
ManualData<-read_excel("FigS1,S2-Sensor-mV.xlsx",sheet = "Sensor-leafmV", col_types = c("text", "numeric","numeric","numeric","numeric","text")) # loading excel, defining which sheet is loaded and defining the type of each column
ManualData <- na.omit(ManualData) #removes all rows with NA - found in some of WM-12 data
ManualData$Cultivar <- factor(ManualData$Cultivar, levels = c("WM-12", "WM-1","G122","Eldorado","ICA Bunsi","Beryl", "Othello","Montrose")) #order the Cultivars (for graph later on)
ManualData$TimePoint <- factor(ManualData$TimePoint, levels = c("Pre treatment", "Post treatment")) #order the treatments pre and post (for graph later on)
```

# Fig. S1 - Leaf mV (Wounding Treatement) and Fig. S2 Leaf mV (Inoculation Treatment)
## Splitting Data
```{r include=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
Mv_WoundData<- ManualData%>% filter(Block == 4) #Create a new table from ManualData that just has Block 4 (wounding) - will be used for FigureS1
Mv_InoculationData<- ManualData %>% filter(Block < 4) #removes Block 4 from ManualData (wounding) - Block 1 -3 were inoculated - this data will be used for Plot Fig. S2
```

## Generating Sup. S1 - Leaf mV (Wounded)
```{r echo=FALSE}
Mv_WoundData$Leaf <-factor(Mv_WoundData$Leaf, levels = c("1", "2", "3")) #defines Leaf number as a factor, needed to ensure that the group below can be grouped by Leaf
FigS1<-ggboxplot(Mv_WoundData,"Cultivar", "mV",facet.by = "TimePoint", color = "Leaf", palette="grey", title="Sensor - Leaf Surface mV") + #facet.by = group plots by TimePoint (pre and post), and group by color = Leaf, palette is greyscale
ylab("Millivolt (mV)") + #renam y axis
xlab("") + #blank out x axis label
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1)) + # x axis font size + change orientation of x axis labels
theme(axis.text.y = element_text(size=12))+ # y axis font size
theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="grey", linetype="blank")) + #adding grid lines to graph, minor grids (0.5) are blank to remove them
theme(plot.title = element_text(hjust = 0.5, vjust=2))+ #Centering title
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5) #adding in a dashed line
FigS1<-ggpar(FigS1, legend = "bottom") # Legend at bottom
FigS1<-FigS1+facet_grid(rows=vars(TimePoint)) #Creates 2 rows of graphs of Pre and Post
FigS1<- annotate_figure(FigS1, top = text_grob("Wounding Treatment", color = "black", face = "bold", size = 14)) #Adding a main title above the panels
FigS1
```

## Saving Fig. S1 as a tiff and pdf
```{r eval=FALSE, include=FALSE}
ggsave(plot=FigS1, file="FigureS1-mV-Wounded-Pre&Post.tiff", units="in", width=7.5, height=7.5, dpi=300,compression='lzw')
ggsave(plot=FigS1, file="FigureS1-mV-Wounded-Pre&Post.pdf", units="in", width=7.5, height=7.5, dpi=300)
```


## Generating Fig. S2 - Leaf mV (Inoculation Treatment)
```{r echo=FALSE}
Mv_InoculationData$Leaf <-factor(Mv_InoculationData$Leaf, levels = c("1", "2", "3", "4", "5", "6", "7", "8", "9")) #defines Leaf number as a factor, needed to ensure that the group below can be grouped by Leaf
FigS2<-ggboxplot(Mv_InoculationData,"Cultivar", "mV",facet.by = "TimePoint", color = "Leaf", palette="grey", title="Sensor - Leaf Surface mV", outlier.size=0.5, legend = "bottom") + #facet.by = group plots by TimePoint (pre and post), and group by color = Leaf, palette is greyscale
ylab("Millivolt (mV)") + #renames y axis
xlab("") + #blank out x axis label
theme(axis.text.x = element_text(size=12,angle = 90,vjust = 0.4, hjust=1)) + # x axis font size + change orientation of x axis labels
theme(axis.text.y = element_text(size=12))+ # y axis font size
theme(panel.grid.major.y = element_line(colour = "grey", linetype="dashed",size=0.5), panel.grid.minor.x = element_line(colour="grey", linetype="blank")) + #adding grid lines to graph, minor grids (0.5) are blank to remove them
theme(plot.title = element_text(hjust = 0.5, vjust=2))+ #Centering title
geom_vline(xintercept = 3.5, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line
geom_vline(xintercept = 6.5, linetype="dashed", color = "black", size=0.5) + #adding in a dashed line
guides(colour = guide_legend(nrow = 1)) #legend below all 1 line instead of 2 lines
FigS2<-FigS2+facet_grid(rows=vars(TimePoint)) #Creates 2 rows of graphs of Pre and Post
FigS2<- annotate_figure(FigS2, top = text_grob("Inoculation Treatment (Wounding + Ss)", color = "black", face = "bold", size = 14)) #Adding a main title above the panels
FigS2
```

## Saving Sup. Fig. 1 as a tiff and pdf
```{r eval=FALSE, include=FALSE}
ggsave(plot=FigS2, file="FigureS2-mV-Inoculation-Pre&Post.tiff", units="in", width=11, height=7.5, dpi=300,compression='lzw')
ggsave(plot=FigS2, file="FigureS2-mV-Inoculation-Pre&Post.pdf", units="in", width=11, height=7.5, dpi=300)
```
